Higher-Order unification for free!

### 2024-04-11 paper/main.tex
link-beta -> link-llam
other rename beta to llam

\newcommand{\linkbeta}{\linkMacro{\beta}}
\newcommand{\linketa} {\linkMacro{\eta}}

@@

\newcommand{\linketaM}[3]{\ensuremath{#1 \vdash #2 =_\eta #3}}
\newcommand{\linkbetaM}[3]{\ensuremath{#1 \vdash #2 =_\beta #3}}
\newcommand{\substCell}[3]{\ensuremath{#1 \vdash #2 = #3}}
\newcommand{\mapping}[3]{\ensuremath{#1 \mapsto #2^#3}}

@@

\input{code/comp_links}
\todo{rename link beta to link llam?}

@@

When detailing examples we write links as equations between two
terms under a context.
The equality sign is subscripted with
kind of \elpiIn{baselink}. For example $\linkbetaM{x}{A_x}{F_x~a}$ corresponds to:
\begin{elpicode}
abs x\ val (link-beta (uva A [x]) (app[uva F [x],con "a"]))
\end{elpicode}

@@

Given a list of unification problems, $\foUnifPb_1\dots
\foUnifPb_n$ with $\foUnifPb_n$ in \notllambda, it is often the case that the resolution of
$\bigwedge_{i=0}^{n-1}\foUnifPb_i$ gives a partial substitution $\rho$, such
that $\rho\foUnifPb_n$ falls again in \llambda.
%
% make test ONLY=7003 TEX=TEX + deactivate beta tex
\printAlll
  {{{X,\lambda x.Y},
    {(X\appsep a),a}}}
  {{{A,\lambda x.B},
    {(A\appsep a),a}}}
  {{{Y,B,0},
    {X,A,0}}}
  {{}}

In the example above, we see that $\foUnifPb_1$ instantiates $X$ so that
$\foUnifPb_2$ can be solved in \llambda.
% \noteE[inline]{it is even a ground term, there is no unification left to perform actually}
% \noteD[inline]{i don't understand the note}
\noteD{Cambiare esempio riprendendo quello di enrico}
On the other hand, we see that, 
\Ue can't solve the compiled problems \hoUnifPb. In
fact, the resolution of $\hoUnifPb_1$ gives the substitution $\sigma = \{ A
\mapsto \lambda x. B\}$, but the dereferencing of $\hoUnifPb_2$ gives the 
non-unifiable problem $(\lambda x. B) \appsep a \nUe a$.

@@

\newcommand{\progBetaLL}{\emph{progress-beta-\llambda}\xspace}
\newcommand{\progBetaRH}{\emph{progress-beta-rigid-\rhs}\xspace}
\newcommand{\progBetaDedup}{\emph{progress-beta-dedup}\xspace}
\newcommand{\progBetaLH}{\emph{progress-beta-rigid-\lhs}\xspace}

>>>>

\newcommand{\linkbeta}{\linkMacro{\llambda}}
\newcommand{\linketa} {\linkMacro{\eta}}

@@

\newcommand{\linketaM}[3]{\ensuremath{#1 \vdash #2 =_\eta #3}}
\newcommand{\linkbetaM}[3]{\ensuremath{#1 \vdash #2 =_{\llambda} #3}}
\newcommand{\substCell}[3]{\ensuremath{#1 \vdash #2 = #3}}
\newcommand{\mapping}[3]{\ensuremath{#1 \mapsto #2^#3}}

@@

\input{code/comp_links}

@@

When detailing examples we write links as equations between two
terms under a context.
The equality sign is subscripted with
kind of \elpiIn{baselink}. For example $\linkbetaM{x}{A_x}{F_x~a}$ corresponds to:
\begin{elpicode}
abs x\ val (link-llam (uva A [x]) (app[uva F [x],con "a"]))
\end{elpicode}

@@

Given a list of unification problems, $\foUnifPb_1\dots
\foUnifPb_n$ with $\foUnifPb_n$ in \notllambda, it is often the case that the resolution of
$\bigwedge_{i=0}^{n-1}\foUnifPb_i$ gives a partial substitution $\rho$, such
that $\rho\foUnifPb_n$ falls again in \llambda.
%
% make test ONLY=7003 TEX=TEX + deactivate beta tex
\printAlll
  {{{X,\lambda x.a},
    {(X\appsep a),a}}}
  {{{A,\lambda x.a},
    {(A\appsep a),a}}}
  {{{X,A,0}}}
  {{}}

In the example above, we see that $\foUnifPb_1$ instantiates $X$ so that
$\foUnifPb_2$ can be solved in \llambda.
On the other hand, we see that, 
\Ue can't solve the compiled problems \hoUnifPb. In
fact, the resolution of $\hoUnifPb_1$ gives the substitution $\sigma = \{ A
\mapsto \lambda x. a\}$, but the dereferencing of $\hoUnifPb_2$ gives the 
non-unifiable problem $(\lambda x. a) \appsep a \nUe a$.

@@

\newcommand{\progBetaLL}{\emph{progress-\llambda-refine}\xspace}
\newcommand{\progBetaRH}{\emph{progress-\llambda-\rhs}\xspace}
\newcommand{\progBetaDedup}{\emph{progress-\llambda-dedup}\xspace}
\newcommand{\progBetaLH}{\emph{progress-\llambda-\lhs}\xspace}
